---
title: "High Tide Tax"
author: "Will Godwin"
date: "12/22/2019"
output: pdf_document
---

```{r setup, echo=F, warning=F, message=F}
##########################################################################################################
#setup####################################################################################################
##########################################################################################################
#data downloaded here: https://www.climatecosts2040.org/data-sets
library(tidyverse); library(openxlsx); library(sf); library(biscale); library(cowplot); library(gridGraphics)

#filepaths and filenames
cost_datapath <- paste0(here::here(),"/seawall/data/Climate_Costs_2040_Data/Con Dist Cost Summary.xlsx")
dist_datapath <- paste0(here::here(),"/seawall/data/Climate_Costs_2040_Data/Con Dist Length Summary.xlsx")
shape_filepath <- paste0(here::here(),"/google_ads/data/districtShapes/districts114.shp")
demo_filepath <- paste0(here::here(),"/seawall/data/demographics/ACS_17_5YR_DP03_with_ann.csv")
pop_filepath <- paste0(here::here(),"/seawall/data/population/ACS_17_5YR_DP05_with_ann.csv")
out_dir <- paste0(here::here(),"/seawall/output/")

##########
#Wrangle the cost data
#find sheets that hold data of interest
state_vec <- getSheetNames(cost_datapath)
state_vec <- state_vec[nchar(state_vec)==2] #where the number of characters is 2

#read in all sheets
df <- bind_rows(
  lapply(state_vec, function(x){
    #read in data
    read.xlsx(cost_datapath, sheet = x) %>% 
      mutate(state=x) %>% #generate state variable
      filter(X1=="RCP 4.5, 2040, 50th%, 1 year return")
  })
)

#reshape the data to long format
df <- pivot_longer(df, cols= contains("district"), names_to = "district", values_drop_na = T) %>%
  mutate(district=substr(district, 24, 26)) %>%
  select(state, district, value)
  
##########
#Wrangle the length of wall data
#find sheets that hold data of interest
state_vec <- getSheetNames(dist_datapath)
state_vec <- state_vec[nchar(state_vec)==2] #where the number of characters is 2
state_vec <- setdiff(state_vec, "PA") #no data for PA for some reason...

#read in all sheets
df_dist <- bind_rows(
  lapply(state_vec, function(x){
    #read in data
    t <- read.xlsx(dist_datapath, sheet = x) %>% 
      mutate(state=x) %>% #generate state variable
      filter(X1=="RCP 4.5, 2040, 50th%, 1 year return")
    return(t)
  })
)

#reshape the data to long format
df_dist <- pivot_longer(df_dist, cols= contains("district"), names_to = "district", values_drop_na = T) %>%
  mutate(district=substr(district, 24, 26)) %>%
  select(state, district, dist=value)

##########
#merge sea wall cost and sea wall length data
df <- left_join(df, df_dist, by=c("state", "district")) %>%
  mutate(value_dist=value/dist)

##########################################################################################################
#state postal codes for merging data onto shapefile
states <- cbind(state_abr=state.abb, STATENAME=state.name) %>% as.data.frame()

#read in shapefile of congressional districts
shape <- read_sf(shape_filepath)
  
#subset and merge with cost data
shape_data <- shape %>%
  left_join(states, by=c("STATENAME")) %>% 
  filter(state_abr %in% state_vec) %>%
  select(district=DISTRICT, state=state_abr) %>%
  left_join(df, by=c("district", "state"))

#read in population to generate per capita cost
df_pop <- read.csv(pop_filepath) %>%
  select(contains("GEO"), pop=HC01_VC03) %>%
  mutate(district=as.character(substr(GEO.display.label, 24, 25)),
         district=str_trim(district, side = "right"),
         STATENAME=sub('.*,\\s*','', GEO.display.label)) %>%
  left_join(states, by=c("STATENAME")) %>%
  filter(!is.na(state_abr)) %>%
  select(pop, district, state=state_abr)

#combine with seawall data and standardize by population
shape_data <- left_join(shape_data, df_pop, by=c("district", "state")) %>%
  mutate(pop=as.numeric(as.character(pop)),
         value_percap=value/pop,
         value_percap_cat=cut(value_percap, breaks = c(0,350,1500,4000,10000,40000), 
                              labels = c("0-350", "350-1,500", "1,500-4,000", "4,000-10,000", "10,000+"))) %>%
  filter(!is.na(value_percap)) #some values are missing for some reason-messes up the bivariate classes
```


```{r cost by mile map, echo=F, warning=F}
#plot cost by mile of sea wall
shape_data %>%
  mutate(value_dist=value_dist/10^6) %>%
  filter(!is.na(value_dist)) %>%
  ggplot() + 
  geom_sf(aes(color=value_dist, fill=value_dist)) +
  #scale_fill_viridis_c(name="USD per capita") +
  ggtitle("US Seawall cost per mile-2040 with 1 year surge") +
  scale_fill_viridis_c(name="USD in millions") +
  scale_color_viridis_c(name="USD in millions") +
  theme_bw()
```

Average cost of seawall construction per mile by congressional district. Calculated by dividing the estimated cost of seawall construction by the estimated length of sea wall in miles. All estimates refer to 2040 predictions of RCP 4.6, with 1-year storm surge.

\newpage

```{r cost per capita map, echo=F, warning=F}
#plot seawall cost per capita
shape_data %>%
  mutate(value=value/10^6) %>%
  filter(!is.na(value_percap_cat)) %>%
  ggplot() + 
    geom_sf(aes(fill=value_percap_cat)) +
    #scale_fill_viridis_c(name="USD per capita") +
    ggtitle("US Seawall cost per capita-2040 with 1 year surge") +
    scale_fill_manual(values=c("#f6eff7", "#bdc9e1", "#67a9cf", "#1c9099", "#016c59"), name="USD") +
    scale_color_manual(values=c("#f6eff7", "#bdc9e1", "#67a9cf", "#1c9099", "#016c59"), name="USD") +
    theme_bw()
```

Average cost of seawall construction per capita by congressional district. Population estimates sourced from [American Community Survey in 2017](https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_17_5YR_DP05&src=pt).

\newpage

```{r poverty prep, echo=F, warning=F}
#poverty data: https://factfinder.census.gov/faces/tableservices/
df_pov <- read.csv(demo_filepath) %>%
  select(contains("GEO"), poverty=HC03_VC161) %>% #percent of families/people whose income below poverty in past 12 months-2017
  mutate(district=as.character(substr(GEO.display.label, 24, 25)),
         district=str_trim(district, side = "right"),
         STATENAME=sub('.*,\\s*','', GEO.display.label),
         poverty=as.numeric(as.character(poverty))) %>%
  left_join(states, by=c("STATENAME")) %>%
  filter(!is.na(district)) %>%
  select(poverty, district, state=state_abr)

#combine with seawall data
shape_data <- left_join(shape_data, df_pov, by=c("district", "state"))

#create classes
shape_data <- bi_class(shape_data, x = value_percap, y = poverty, style = "quantile", dim = 3)
```


```{r cost by poverty map, echo=F, warning=F}
#bivariate plot
map <- ggplot() +
  geom_sf(data = shape_data, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = F) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  labs(title = "Seawall cost and poverty") +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
          dim = 3,
          xlab = "Higher seawall costs ",
          ylab = "Higher % in poverty ",
          size = 8) 

ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, x=0.25, y=.35, 0.4, 0.4)
```

Bivariate map illustrating seawall costs per capita and percent of district living at or below the poverty line. The more magenta, the poorer the district and the greener, the most costly seawall per capita. The green and magenta colors converge on dark blue as both poverty and seawall costs increase. Poverty data also from [American Community Survey in 2017](https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_17_5YR_DP05&src=pt).