---
title: "NAITRE Interim Analysis"
author: "Will Godwin & Ben Arnold"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '1'
---


```{r setup, echo=F, message=F, warning=F}
# clear memory
rm(list=ls())

# directories
# You may need to change this start path if you use Box Sync or Windows
box_dir <- "~/Box/"
data_dir <- paste0(box_dir, "NAITRE-unmasked-materials/InterimAnalysis-2020-10/data/")
code_dir <- paste0(box_dir, "NAITRE-unmasked-materials/InterimAnalysis-2020-10/R/")
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

# libraries
library(tidyverse)
library(table1)
library(data.table)
library(epitools)
library(kableExtra)
library(scales)
library(labelled)
library(growthstandards)

# treatment assignments
Az <- c("RR", "SS", "TT", "WW")
Pl <- c("UU", "XX", "YY", "ZZ")
```

```{r read in clean data, echo=F, message=F} 
# refer to near the top of weekly report Rmd for context as to how the data was cleaned and saved
# eligibility
df_eligibility <- read_csv(paste0(data_dir, "eligibility.csv"))

# baseline
df_baseline <- read_csv(paste0(data_dir, "baseline.csv"))

# treatment
df_treatment <- read_csv(paste0(data_dir, "treatment.csv"))

# followup
df_followup <- read_csv(paste0(data_dir, "followup.csv"))

# treatment letters
tr_assign <- read_csv(paste0(data_dir, "treatment_letters.csv"))

# metadata regarding follow-up visits necessary to calculate follow-up coverage
table.followup <- fread(paste0(data_dir, "followup_table.csv"))
```

```{r join in treatment assignment, echo=F, message=F} 
# treatment assignment for eligibility
df_eligibility <- left_join(df_eligibility, tr_assign, by="childID") %>%
  mutate(TL = case_when(
    assignment %in% Az ~ "Azithromycin",
    assignment %in% Pl ~ "Placebo")) %>% 
  as.data.table()

# treatment assignment for baseline
df_baseline <- left_join(df_baseline, tr_assign, by="childID") %>%
  mutate(TL = case_when(
    assignment %in% Az ~ "Azithromycin",
    assignment %in% Pl ~ "Placebo")) %>% 
  as.data.table()

# treatment assignment for treatment form
df_treatment <- left_join(df_treatment, tr_assign, by="childID") %>%
  mutate(TL = case_when(
    assignment %in% Az ~ "Azithromycin",
    assignment %in% Pl ~ "Placebo")) %>% 
  as.data.table()

# treatment assignment for followup form
df_followup <- left_join(df_followup, tr_assign, by="childID") %>%
  mutate(TL = case_when(
    assignment %in% Az ~ "Azithromycin",
    assignment %in% Pl ~ "Placebo")) %>% 
  as.data.table()
```

```{r eligibility criteria, echo=F}
#create screened, eligible, enrolled indicators
df_eligibility[!duplicated(childID) | is.na(childID), screened := 1]
df_eligibility[startDate>as.Date("2019-08-23") & #date we expanded CSPSs hierarchy
                startDate-childBirthdate > 7 &
                 startDate-childBirthdate < 28 &
                 feedOrally == 1 &
                 stay6months == 1 & 
                 allergyAzalides == 2 &
                 village != 103 &
                 (childWeight == 1 | is.na(childWeight)) &
                 ((district %in% c(1,2,4,5,6,7,8,10) & (inrange(village, 1, 102) | 
                                                          village==999 | 
                                                          inrange(village, 104, 250))) |
                    (district==3 & otherStudy==2 & (inrange(village, 1, 102) | 
                                                      inrange(village, 104, 250)))) &
                 (inrange(region, 1, 5) | is.na(region)) &
                 (inrange(q_2_1_6, 1, 30) | inrange(q_2_1_6, 32, 45)) &
                 screened == 1,
               eligible := 1]

# slightly different eligibility criteria since we expanded total CSPSs
df_eligibility[startDate<=as.Date("2019-08-23") &
               startDate-childBirthdate > 7 &
               startDate-childBirthdate < 28 &
                 feedOrally == 1 &
                 stay6months == 1 & 
                 allergyAzalides == 2 &
                 otherStudy == 2 &
                 !is.na(village) & village != 103 &
                 (childWeight == 1 | is.na(childWeight)) &
                 screened == 1,
               eligible := 1]
df_eligibility[eligible == 1 & consent == 1 & !is.na(childID), enrolled := 1]
df_eligibility[(consent == 2 | (is.na(consent) & eligible == 1) | 
                 is.na(childID)) & eligible == 1,
               refused := 1]

```

```{r prep consort, echo=F, message=F, warning=F}
# screened
screened <- df_eligibility %>% summarise(sum(screened, na.rm = T)) %>% pull

# total eligible
eli <- df_eligibility %>% summarise(sum(eligible, na.rm = T)) %>% pull

# ineligible
ineli <- screened - eli

# total enrolled and randomized
enrolled <- df_eligibility %>% summarise(sum(enrolled, na.rm = T)) %>% pull

# refused entry to study
not_enr <- eli - enrolled

# treatment numbers by arm
az_N <- df_treatment %>% filter(TL=="Azithromycin" & isAdministered==1) %>% distinct(childID) %>% nrow
pl_N <- df_treatment %>% filter(TL=="Placebo" & isAdministered==1) %>% distinct(childID) %>% nrow

# reached 6-mo timepoint or died
az_6mo <- df_followup %>% filter(TL=="Azithromycin" & (timepoint %in% c(180, 365) | vitalStatus == 2)) %>%
  distinct(childID) %>% nrow
pl_6mo <- df_followup %>% filter(TL=="Placebo" & (timepoint %in% c(180, 365) | vitalStatus == 2)) %>%
  distinct(childID) %>% nrow
```

```{r get the lost to follow-up numbers, echo=F}
# create a copy of follow-up with new name that the script below expects
dt.followup <- copy(df_followup)
dt.eligibility <- copy(df_eligibility)

# Important!! Set this date to the date of data cutoff for the analysis. 
# The follow-up calculate relies on it
date_cutoff <- "2020-09-28"

# script that loops through each follow-up timepoint (21 day, 3-mo, 6-mo, etc.) and calculates coverage
source(paste0(code_dir, "calc_followup_by_arm.R"), local = T)

# calculate lost to followup for AZ arm
az_ltfu <- counts.all %>% 
  filter(timepoint=="6-month" & TL=="Azithromycin") %>% 
  summarise(sum(expected) - (sum(actual) + sum(window))) %>% pull

# calculate lost to followup for PL arm
pl_ltfu <- counts.all %>% 
  filter(timepoint=="6-month" & TL=="Placebo") %>% 
  summarise(sum(expected) - (sum(actual) + sum(window))) %>% pull
```

## CONSORT (Display 1)

```{r render consort diagram, echo=F, message=F, warning=F}
# make diagram foundation
data <- tibble(x=1:100, y=1:100)
p <- data %>% 
  ggplot(aes(x, y)) +
  scale_x_continuous(minor_breaks = seq(10, 100, 10)) +
  scale_y_continuous(minor_breaks = seq(10, 200, 10)) +
  theme_linedraw()

# size of font
s <- 2.5

# draw boxes
#consort <- 
  p +
  # top box
  geom_rect(xmin=36, xmax=64, ymin=92, ymax=100, color='black',
            fill='white', size=1) +
  annotate('text', x=50, y=96, label=paste0("Total screened - ", screened), size=s) +
  
  # first arrow
  geom_segment(x=50, xend=50, y=92, yend=81, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
  
  # second box
  geom_rect(xmin=36, xmax=64, ymin=72, ymax=80, color='black',
            fill='white', size=1) +
  annotate('text', x=50, y=76, label=paste0("Total eligible - ", eli), size=s) +
  
  # second arrow
  geom_segment(x=50, xend=50, y=72, yend=61, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +

  # Third box
  geom_rect(xmin=36, xmax=64, ymin=52, ymax=60, color='black',
            fill='white', size=1) +
  annotate('text', x=50, y=56, label=paste0("Total enrolled & randomized - ", enrolled), size=s) +
  
  # arrow to exclude box1
  geom_segment(x=50, xend=69, y=86, yend=86, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
    
  #exclude box1
  geom_rect(xmin=70, xmax=96, ymin=82, ymax=90, color='black',
            fill='white', size=1) +
  annotate('text', x=83, y=86, label=paste0("Total ineligible - ", ineli), size=s) +

  # arrow to exclude box2
  geom_segment(x=50, xend=69, y=66, yend=66, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
    
  #exclude box2
  geom_rect(xmin=70, xmax=96, ymin=62, ymax=70, color='black',
            fill='white', size=1) +
  annotate('text', x=83, y=66, label=paste0("Total refused - ", not_enr), size=s) +
  
  # first left arrow
  geom_segment(x=36, xend=19, y=56, yend=56, 
    size=1, linejoin = "mitre", lineend = "butt") +
  geom_segment(x=19, xend=19, y=56, yend=53, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
  
  # First left box
  geom_rect(xmin=6, xmax=32, ymin=44, ymax=52, color='black',
            fill='white', size=1) +
  annotate('text', x=19, y=48, label=paste0("Azithromycin treated - ", az_N), size=s) +
  
  # second left arrow
  geom_segment(x=19, xend=19, y=44, yend=37, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
  
  # first right arrow
  geom_segment(x=64, xend=81, y=56, yend=56, 
    size=1, linejoin = "mitre", lineend = "butt") +
  geom_segment(x=81, xend=81, y=56, yend=53, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
    
  # First right box
  geom_rect(xmin=68, xmax=94, ymin=44, ymax=52, color='black',
            fill='white', size=1) +
  annotate('text', x=81, y=48, label=paste0("Placebo treated - ", pl_N), size=s) +

  # second right arrow
  geom_segment(x=81, xend=81, y=44, yend=37, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
  
  # Second left box
  geom_rect(xmin=6, xmax=32, ymin=28, ymax=36, color='black',
            fill='white', size=1) +
  annotate('text', x=19, y=32, label=paste0("Azithromycin lost to followup - ", az_ltfu), size=s) +
  
  # third left arrow
  geom_segment(x=19, xend=19, y=28, yend=21, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
    
  # Second right box
  geom_rect(xmin=68, xmax=94, ymin=28, ymax=36, color='black',
            fill='white', size=1) +
  annotate('text', x=81, y=32, label=paste0("Placebo lost to followup - ", pl_ltfu), size=s) +
  
  # third right arrow
  geom_segment(x=81, xend=81, y=28, yend=21, 
    size=1, linejoin = "mitre", lineend = "butt",
    arrow = arrow(length = unit(1, "mm"), type= "closed")) +
    
  # third left box
  geom_rect(xmin=6, xmax=32, ymin=12, ymax=20, color='black',
            fill='white', size=1) +
  annotate('text', x=19, y=16, label=paste0("Azithromycin reached 6-mo - ", az_6mo), size=s) +
  
  # third right box
  geom_rect(xmin=68, xmax=94, ymin=12, ymax=20, color='black',
            fill='white', size=1) +
  annotate('text', x=81, y=16, label=paste0("Placebo reached 6-mo - ", pl_6mo), size=s) +
  theme_void()
```

## Enrollment characteristics (Display 2)
```{r baseline indicators, echo=F, message=F, warning=F}
# custom render function for continuous variables - just show MEAN and SD
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (%s)", MEAN, SD)))
}

# calculate the waz, haz, and whz scores using the "growthstandards" package
df_baseline <- df_baseline %>%
  mutate(sex=if_else(q_2_1_1==1, "Male", "Female"),
         ageInDays=as.numeric(difftime(startDate, childBirthdate, units = "days")),
         haz_zscore=who_value2zscore(ageInDays, length, x_var = "agedays", y_var = "htcm", sex = sex),
         waz_zscore=who_value2zscore(ageInDays, weight, x_var = "agedays", y_var = "wtkg", sex = sex),
         whz_zscore=who_value2zscore(length, weight, x_var = "lencm", y_var = "wtkg", sex = sex))

# prepare variable for rendering table
df_baseline %>%
  mutate(Female = ifelse(q_2_1_1==2, "Female", "Male"),
         age_cat = 
           case_when(
             ageInDays %in% seq(8, 13) ~ "8-13d",
             ageInDays %in% seq(14, 20) ~ "14-20d",
             ageInDays %in% seq(21, 28) ~ "21-28d"
           ),
         birthweight = ifelse(q_2_1_2>0, q_2_1_2, NA)) %>%
  select(Female, ageInDays, age_cat, Birthweight=birthweight, Weight=weight, Height=length,
         WAZ=waz_zscore, HAZ=haz_zscore, WHZ=whz_zscore, TL) %>%
  mutate(age_cat = factor(age_cat, levels = c("8-13d", "14-20d", "21-28d"))) %>%
  set_variable_labels(Female = "Sex", ageInDays = "Age (days)", age_cat = "Age bin",
                      Birthweight = "Birthweight (grams)", Weight = "Weight (kg)",
                      Height = "Height (cm)") %>%
  table1(~ Female + ageInDays + age_cat + Birthweight + Weight + 
           Height + WAZ + HAZ + WHZ | TL, data = .,
         render.continuous=my.render.cont, render.missing=NULL)
```

Birthweight based on self-report from mother/guardian.

## Treatment coverage (Display 3)
```{r treatment coverage, echo=F, message=F}
# special render function to only produce frequency (not percentage too) 
my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d", FREQ))))
}

#treatment coverage
bind_rows(
  # total enrolled
  df_eligibility %>%
    filter(enrolled==1) %>%
    select(childID, TL) %>%
    mutate(Status = "Enrolled"),
  
  # total treated
  df_treatment %>%
    # find unique participants to avoid counting duplicates twice
    filter(isAdministered==1 & !duplicated(childID)) %>%
    select(childID, TL) %>%
    mutate(Status = "Treated")

  # # link by childID
  # by="childID"
) %>%
  table1(~ Status | TL, data = ., overall = F,
         render.categorical=my.render.cat)
```

Counts of study participants that were enrolled and treated, as well as the percent of the enrolled that were treated.

## Follow-up visits (Display 4)
```{r followup coverage, echo=F}
#Prepare table
counts.all[, timepoint := fct_relevel(timepoint, c("7-day", "14-day", "21-day",
                                                   "7-21 day", "3-month", "6-month",
                                                   "12-month"))]
counts.overall <- counts.all[,.(expected = sum(expected), actual = sum(actual),
                                window = sum(window)), by=c("timepoint", "TL")]

#add overall row
row.overall <- data.table(timepoint="Overall", 
                          counts.all[,.(expected = sum(expected), actual = sum(actual),
                                        window = sum(window)), by="TL"])
counts.overall <- rbind(counts.overall, row.overall)
counts.overall[, Ratio := percent(actual/expected, accuracy = .1)]

#table
setorderv(counts.overall, c("timepoint", "TL"))
bold_rows <- which(counts.overall[,timepoint]=="Overall")
kable(counts.overall, align = "c", col.names = c("Timepoint", "Arm", "Expected visits",
                                                 "Actual visits", "In window", "Follow-up (%)")) %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = TRUE) %>% 
  row_spec(c(0, bold_rows), bold=TRUE)
```
*In window* refers to number of participants missing follow-up but still within the respective time window (+3 days for 7-day, 14-day, and 21-day follow-up; +21 days for 3-month; +42 days for 6-month and 12-month follow-up). *Expected visits* for each child are based on the days between enrollment date and the report date. For example, a child enrolled 10 days ago would have an expected 7-day visit but not an expected 14-day visit.
\newline **Timepoints**: 7, 14, and 21 days *after treatment*. 3, 6, and 12 months *after birth*. We are currently only conducting 1 interview within 21 days of treatment - shown in the 7-21 day timepoint.

## Adverse events (Display 5)
```{r ae, echo=F}
#adverse event counts by arm
df_followup %>%
  # q_5_6__7 column represents "none of the above" for adverse events
  select(childID, TL, timepoint, starts_with("q_5_6__"), -q_5_6__7) %>%
  gather(key="key", value = "value",
         starts_with("q_5_6__")) %>%
  filter(value==1) %>%
  mutate(value2 = case_when(
    key == "q_5_6__1" ~ "Fever",
    key == "q_5_6__2" ~ "Diarrhea",
    key == "q_5_6__3" ~ "Vomiting",
    key == "q_5_6__4" ~ "Abdominal pain",
    key == "q_5_6__5" ~ "Rash",
    key == "q_5_6__6" ~ "Constipation"
  )) %>%
  set_variable_labels(value2 = "Adverse Events") %>%
  table1(~ value2 | TL, data = .)
```

Each participant was interviewed at least once in 7-21 days following treatment. The display above shows the number of adverse events reported in the 7-21 days following treatment.

## Hospitalization and/or death (Display 6)
```{r hospitalizations and deaths, echo=FALSE}
#counts by region of children that have died or been hospitalized
df_followup %>%
  filter(hospitalVisit==1 | vitalStatus==2) %>%
  mutate(hosp_o_death = "Total") %>%
  set_variable_labels(hosp_o_death = "Hospitalization and/or death") %>%
  table1(~ hosp_o_death | TL, data = ., render.categorical=my.render.cat)
```

Total participants hospitalized and/or died as reported at 7-21 day, 3-month, 6-month, and 12-month visits.

```{r calculate person time, echo=F}
#calculate person-time at risk
#find most recent follow-up visit for each child
df_followup[, recent_timepoint := max(timepoint), by="childID"]
df_followup[, recent_timepoint6mo := ifelse(recent_timepoint==365, 180, recent_timepoint)]
df_followup[recent_timepoint != timepoint, sec_recent_timepoint := max(timepoint), by="childID"]
dt.persont <- df_followup[timepoint==recent_timepoint | timepoint==sec_recent_timepoint,
                          .(childID, startDate, vitalStatus, timepoint, region_name, 
                            recent_timepoint, sec_recent_timepoint, recent_timepoint6mo, TL)]
dt.persont <- merge(dt.persont, df_baseline[,.(childID, startDate)], by="childID", all.x=T)

#subtract the number of days between enrollment and follow-up visit
died_6mo_ids <- df_followup %>% filter(vitalStatus==2 & timepoint<=180) %>% pull(unique(childID))
died_ids <- df_followup %>% filter(vitalStatus==2) %>% pull(unique(childID))

#calculate 6-mo person time
dt.persont_6mo <- bind_rows(
  #the children that have not died
  dt.persont %>%
    filter(childID %in% setdiff(childID, died_6mo_ids) & 
             timepoint==recent_timepoint6mo & timepoint<=180) %>%
    mutate(person_days=as.numeric(difftime(startDate.x, startDate.y, units = "days"))),
  #the children that have died
  bind_rows(
    
    #find the person time from baseline to visit before death visit
    dt.persont %>%
      filter(childID %in% died_6mo_ids & (vitalStatus==1 & 
                                            timepoint==sec_recent_timepoint & timepoint<=90) 
                                      | (vitalStatus==2 & timepoint==7)) %>%
      mutate(person_days=as.integer(difftime(startDate.x, startDate.y, units = "days"))),
    
    #find the midpoint between final alive visit and death interview
    dt.persont %>%
      filter(childID %in% died_6mo_ids & (timepoint==recent_timepoint6mo | 
                                        timepoint==sec_recent_timepoint)) %>%
      arrange(childID, -vitalStatus) %>%
      group_by(childID, TL) %>%
      summarise(person_days=as.integer(difftime(first(startDate.x), last(startDate.x)))/2)
  ) %>% 
    group_by(childID, TL) %>%
    summarise(person_days=sum(person_days))
) %>% 
  data.table()
```

## Six-month mortality (Display 7)
```{r deaths at 6 months, echo=FALSE}
#number of deaths by arm and overall
n_died6mo <- df_followup %>%
  left_join(df_baseline %>% select(childID, enrollDate=startDate), by="childID") %>% 
  ##all deaths within 8 months of enrollment are included in 6-month mortality
  filter(vitalStatus==2 & difftime(startDate, enrollDate, units = "days")<245) %>%
  group_by(TL) %>%
  summarise(Deaths=n()) %>% as.data.table()

n_died6mo.tot <- df_followup %>%
  left_join(df_baseline %>% select(childID, enrollDate=startDate), by="childID") %>% 
   ##all deaths within 8 months of enrollment are included in 6-month mortality
  filter(vitalStatus==2 & difftime(startDate, enrollDate, units = "days")<245) %>%
  summarise(Deaths=n()) %>% mutate(TL="Overall") %>% as.data.table()

#bind together
n_died6mo <- rbind(n_died6mo, n_died6mo.tot)

#person time by city and overall
dt.persont.tab <- dt.persont_6mo[,.(person_time = round(sum(person_days, na.rm = T)/365,2)),
                             by="TL"]
dt.persont.tot <- dt.persont_6mo[,.(person_time = round(sum(person_days, na.rm = T)/365,2),
                                TL="Overall")]
dt.persont.tab <- rbind(dt.persont.tab, dt.persont.tot)

#merge together deaths and person time to calculate rate
death6rate <- merge(n_died6mo, dt.persont.tab, by="TL", all.y = T)

#estimate rate and CI's using fxn from epitools package
death6rate[, rate := round(pois.exact(Deaths, person_time)[[3]]*1000, 1)]
death6rate[, lower := round(pois.exact(Deaths, person_time)[[4]]*1000, 1)]
death6rate[, upper := round(pois.exact(Deaths, person_time)[[5]]*1000, 1)]
death6rate[, rate := paste0(rate, " (", lower, " to ", upper, ")")]
death6rate[, c("lower", "upper") := NULL]

## make table
death6rate %>%
  arrange(Deaths) %>%
  kable(col.names = c("Arm", "Deaths", "Person-years", 
                                "Mortality/1000 py (95% CI)")) %>%
  row_spec(c(0, 3),bold=TRUE) %>%
  kable_styling() %>%
  column_spec(column = c(1,2,3,4), width = "150px") 
```

## Vital status at 6-months (Display 8)
```{r vital status, echo=F}
# join with baseline form to link enroll date with follow-up date
df_followup %>%
  left_join(df_baseline %>% select(childID, enrollDate=startDate),
            by="childID") %>%
  # subset to participants at 6-mo visit or participants that died within 8 months of followup
  filter(timepoint == 180 | (vitalStatus==2 & difftime(startDate, enrollDate, units = "days")<245)) %>%
  mutate(
    status_clean =
      case_when(
        vitalStatus == 1 ~ "Alive",
        vitalStatus == 2 ~ "Dead",
        vitalStatus == 3 ~ "Moved",
        vitalStatus == 4 ~ "Unknown"
      )
  ) %>% 
  set_variable_labels(status_clean = "Vital Status") %>%
  table1(~ status_clean | TL, data = .)
```

## Hazard ratio permutation distribution (Display 9)
```{r glm model, echo=F}
# prep the data - include all participants who have reached 6-mo visit or died before it
model_df <- df_followup %>%
  left_join(df_baseline %>% select(childID, enrollDate=startDate), by="childID") %>%
  # subset to participants at 6-mo visit or participants that died within 8 months of followup
  filter(timepoint == 180 | (vitalStatus==2 & difftime(startDate, enrollDate, units = "days")<245)) %>%
  mutate(death_response =
           case_when(
             vitalStatus == 1 ~ 0,
             vitalStatus == 2 ~ 1
            )
         )

# set seed
set.seed(2^3)

# run the model
model_df$TL <- factor(model_df$TL, levels = c("Placebo", "Azithromycin"))
mod_empir <- glm(death_response ~ TL, data = model_df, family = binomial(link = cloglog))

# pull the effects
coef_empir <- summary(mod_empir)$coef[2,1]
sd_empir <- summary(mod_empir)$coef[2,2]

# calculate the upper and lower limits on the HR ratio (coefficient from the model)
upper_empir <- exp(coef_empir+(1.96*sd_empir)) %>% round(3)
lower_empir <- exp(coef_empir-(1.96*sd_empir)) %>% round(3)
coef_empir_exp <- exp(coef_empir) %>% round(3)

# permute the treatment allocation and rerun the model to 
# estimate distribution of the model coefficient under the Null
response_vec <- model_df$death_response
treatment_alloc <- model_df$TL

# function that iterates i times, each time saving the model coefficient
permute_fxn <- function(i){
  # pull random numbers the length of treatment vector from uniform distribution 0-1
  reshuffle_vec <- runif(length(treatment_alloc))
  
  # reshuffle the treatment allocation based on random numbers generated
  shuffled_arms <- treatment_alloc[order(reshuffle_vec)]
  
  # run the model on shuffled allocation
  mod <- glm(response_vec ~ shuffled_arms, family = binomial(link = cloglog))
  
  # extract and save the coefficient
  summary(mod)$coef[2,1]
}

# run the function N times to produce the null distribution
N <- 10000
null_vec <- map_dbl(1:N, permute_fxn)

# calculate the permuation test p-value
p_value <- sum(abs(null_vec) > abs(coef_empir)) / length(null_vec)
mod_summary <- summary(mod_empir)
# Summary of model: `r mod_summary`

# plot the null distribution with the empirical value
df <- data.frame(null_vec=null_vec)
ggplot(df, aes(x=null_vec)) + 
  geom_histogram(alpha=0.5, color=NA) +
  geom_vline(xintercept = coef_empir, color="black") +
  xlab("Null distribution") +
  theme_bw()
```

Distribution of the log hazard ratio under the null hypothesis using permutation test. Black line represents the estimated log hazard ratio from data.

Permutation test p-value: `r p_value`

## Mortality test (Display 10)
```{r prepare vital status table, echo=F}
# join with baseline form to link enroll date with follow-up date
df <- df_followup %>%
  left_join(df_baseline %>% select(childID, enrollDate=startDate),
            by="childID") %>%
  # subset to participants at 6-mo visit or participants that died within 8 months of followup
  filter(timepoint == 180 | (vitalStatus==2 & difftime(startDate, enrollDate, units = "days")<245)) %>%
  group_by(TL) %>%
  # tabulate deaths and total participants by arm
  summarise(N_ch=n(), N_d=sum(vitalStatus==2)) %>%
  # calculate cumulative mortality
  mutate(d_percent=percent(N_d/N_ch, accuracy = .01)) %>%
  pivot_wider(names_from=TL, values_from=c("N_ch", "N_d", "d_percent")) %>%
  select(N_ch_Azithromycin, N_d_Azithromycin, d_percent_Azithromycin, 
         N_ch_Placebo, N_d_Placebo, d_percent_Placebo)

# prepare the hazard ratio from the model and p-value
df_hr <- data.frame(hr=paste0(coef_empir_exp, " (", lower_empir, " - ", upper_empir, ")"), pvalue=p_value)

# bind everything together
df <- bind_cols(data.frame(outcome="Death by 6 month visit"), df, df_hr)

# render table
kable(df, col.names = c("Outcome", "N Children", "N Deaths", "%", "N Children",
                     "N Deaths", "%", "HR (95%)", "P-value*"),
      caption = "Mortality difference by arm estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling() %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3,  "Placebo" = 3, " " = 2))
```
